---
title: "EDA_and_stuff_[sam]"
output: html_document
---
```{r, message=FALSE, echo=FALSE}
library(dplyr)
library(pander)
library(ggplot2)
library(tidyr)
```

```{r, message=FALSE, echo=FALSE}
df <- read.csv("games_march2025_cleaned.csv")
keep_col <- c(
  "name","release_date","price","dlc_count","about_the_game","short_description","windows","mac","linux",
  "achievements", "supported_languages","publishers","categories","genres","estimated_owners",
  "average_playtime_forever","peak_ccu","num_reviews_total","pct_pos_total"
)

df2 <- df %>%
  select(any_of(keep_col))
```

```{r Release year, fig.cap = "Figure 3: Number of games released on Steam per year, 1997-2025", echo=FALSE}
df3 <- df2 %>%
  mutate(
    release_date = as.Date(release_date, format="%Y-%m-%d"),
    release_year = as.integer(format(release_date, "%Y"))
  )

df3 %>%
  count(release_year) %>%
  ggplot(aes(x = release_year, y = n)) +
  geom_col() +
  labs(title = "Number of games released on Steam per year, 1997-2025",
       x = "Release year", y = "Number of games") +
  theme_minimal()
```

The number of games released on Steam has followed an increasing trend, which is aligned with the migration away from console gaming towards PC. Note that the data for 2025 is low because it was collected in March 2025 and therefore do not capture the full year.

```{r Pct missing, fig.cap = "Table 2: Percentage of variables missing", message=FALSE, echo=FALSE}
missing_df <- data.frame(
  variable = c("supported_languages: []", "publishers: []", "categories: []", "genres: []",
               "estimated_owners: 0 - 0", "average_playtime_forever: 0", "peak_ccu: 0",
               "num_reviews_total: < 0",
               "pct_pos_total: < 0"),
  pct_missing = c(
    mean(df2$supported_languages == "[]"),
    mean(df2$publishers == "[]"),
    mean(df2$categories == "[]"),
    mean(df2$genres == "[]"),
    mean(df2$estimated_owners == "0 - 0"),
    mean(df2$average_playtime_forever == 0),
    mean(df2$peak_ccu == 0),
    mean(df2$num_reviews_total < 0),
    mean(df2$pct_pos_total < 0)
  ) * 100
)

pander(missing_df)
```

The proportion of data missing as per the above table is most likely due to the way data is collected. They seemed to have used sentinel figures as opposed to NA, e.g. negative or zero values.

```{r N languages & N platforms, message=FALSE, echo=FALSE}
df3 <- df3 %>%
  mutate(
    languages_clean = gsub("\\[|\\]|'", "", supported_languages),
    languages_list  = strsplit(languages_clean, ",\\s*"),
    n_languages     = lengths(languages_list)
  )

df3 <- df3 %>%
  mutate(
    windows = as.logical(windows),
    mac = as.logical(mac),
    linux = as.logical(linux),
    n_platforms = windows + mac + linux
  )
```

```{r Platform availability, fig.cap = "Figure 4: Graph of platform availability by release year, 2005-2025", message=FALSE, echo=FALSE}
platform_by_year <- df3 %>%
  filter(release_year >= 2005) %>%
  group_by(release_year) %>%
  summarise(
    windows = mean(windows) * 100,
    mac     = mean(mac) * 100,
    linux   = mean(linux) * 100
  ) %>%
  pivot_longer(cols = c(windows, mac, linux),
               names_to = "platform", values_to = "pct")

ggplot(platform_by_year, aes(x = release_year, y = pct,
                              color = platform, linetype = platform)) +
  geom_line(linewidth = 1) +
  labs(title = "Graph of platform availability by release year, 2005-2025",
       x = "Release year", y = "Percentage of games released",
       color = "Platform", linetype = "Platform") +
  theme_minimal()
```

The graph above shows the percentage of games' supported platforms released each year. The period prior to 2005 was excluded due to small amount of data that skews the percentage. We can see that almost all games are always available on windows, while linux and mac support have been slowly declining after its peak in 2013.

```{r Est owners, echo=FALSE, message=FALSE, warning=FALSE, fig.cap = "Figure 5: Distribution of estimated owners"}
owner_levels <- c(
  "0 - 0", "0 - 20000", "20000 - 50000", "50000 - 100000",
  "100000 - 200000", "200000 - 500000", "500000 - 1000000",
  "1000000 - 2000000", "2000000 - 5000000", "5000000 - 10000000",
  "10000000 - 20000000", "20000000 - 50000000", "50000000 - 100000000",
  "100000000 - 200000000", "200000000 - 500000000"
)

df3 <- df2 %>%
  count(estimated_owners, name = "Frequency") %>%
  mutate(estimated_owners = factor(estimated_owners,
                       levels = owner_levels,
                       ordered = TRUE)) %>%
  arrange(estimated_owners)

ggplot(df3, aes(x=estimated_owners, y=Frequency))+
  geom_col() +
  labs(title = "Distribution of estimated owners",
       x = "Estimated owners",
       y = "Frequency") +
  theme_minimal() +
  geom_density(fill = "darkslategrey") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r echo=FALSE, fig.cap = "Table 3: Table of two games with highest estimated owners"}
name_estimated_owners <- c("name","estimated_owners")

df_estimated_owners <- data.frame(
  estimated_owners = c(df2[df2$estimated_owners=="100000000 - 200000000", "estimated_owners"],
                       df2[df2$estimated_owners=="200000000 - 500000000", "estimated_owners"]
                       ),
  name = c(df2[df2$estimated_owners=="100000000 - 200000000", "name"],
            df2[df2$estimated_owners=="200000000 - 500000000", "name"]
            )
)

pander(df_estimated_owners)

df3[df3$estimated_owners=="0 - 20000","Frequency"]/sum(df3$Frequency)
```

From the above plot and table, we can see that the estimated_owners distribution is extremely skewed to the right, with only two games holding the highest two values. Approximately 66.19% of games have estimated owners of 0 - 20000.

# proportion of peak_ccu < 1,000
```{r peak ccu, echo=FALSE}
peak_ccu <- df2$peak_ccu[df2$peak_ccu>0]
sum(peak_ccu < 1000)/length(peak_ccu)
```

97.06% of games have peak concurrent users of less than 1,000.

```{r Release year, fig.cap = "Figure 3: Number of games released on Steam per year, 1997-2025", echo=FALSE, results='asis'}
df3 <- df2 %>%
  mutate(
    release_date = as.Date(release_date, format="%Y-%m-%d"),
    release_year = as.integer(format(release_date, "%Y"))
  )

cat("<center>")

df3 %>%
  count(release_year) %>%
  ggplot(aes(x = release_year, y = n)) +
  geom_col() +
  labs(title = "Number of games released on Steam per year, 1997-2025",
       x = "Release year", 
       y = "Number of games") +
  theme_minimal()

cat("</center>")
```
\
The number of games released on Steam has increased significantly, from 1,525 titles in 2014 to 18,282 in 2024, with 2024 alone representing a 43.6% increase on the previous year. The trend is not entirely monotonic, with a decline of 17.4% in 2019. Notably, 67.6% of all games in the dataset were released from 2020 onwards, meaning the catalogue is heavily weighted towards recent titles. This growth is consistent with the lowering of barriers to publication on the platform and the increasing accessibility of game development tools, although the dataset alone cannot distinguish between such explanations. Note that the 2025 figure is low, as the data were collected in March 2025 and therefore do not capture the full year.