---
title: "EDA_graphs"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the page for EDA and graphs. 

* Several summary tables and/or plots, each describing one, two or three variables in the data that you thought were interesting

* Explain the definitions of the variables in each table/plot

* Comment on the main features of each plot

* Include suitable labels and keys for each plot

* Make sure all plots would be readable if printed in black & white, and adjust the point sizes and/or line thicknesses to improve readability

* Lay out all tables so that they are clearly readable and clearly labelled, and do not use excessive significant figures

* All figures and tables should be numbered and should have captions


```{r, echo=FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(data.table)
library(pander)
library(corrplot)
library(pander)


march_c <- read.csv("games_march2025_cleaned.csv")
small_steam <- march_c[c("name","price","dlc_count","about_the_game","short_description","windows","mac","linux",
                         "achievements","supported_languages","publishers","categories","genres",
                         "estimated_owners","average_playtime_forever","peak_ccu","num_reviews_total","pct_pos_total")]
```
\

## Correlation matrix for our selected variables:

```{r, echo=FALSE}
library(corrplot)

numeric_steam <- small_steam[, sapply(small_steam, is.numeric)]
#summary(numeric_steam)

m <- cor(numeric_steam)
#corrplot(m, method = "number")
```

```{r, fig.show="hold", out.width="50%", echo=FALSE}
corrplot(m, method = "number")

pairs(numeric_steam[, 1:7], panel = function(x,y){
  points(x,y)
  abline(lm(y~x), col="red")})
```
```{r}
#install.packages("GGally")
library(GGally)

ggpairs(
  numeric_steam,
  mapping = NULL,
  columns = 1:ncol(numeric_steam),
  title = NULL,
  upper = list(continuous = "cor", 
               combo = "box_no_facet", 
               discrete = "count", 
               na = "na"),
  lower = list(continuous = "points", 
               combo = "facethist", 
               discrete = "facetbar", 
               na = "na"),
  diag = list(continuous = "densityDiag", 
              discrete = "barDiag", 
              na = "naDiag"),
  
  #params = deprecated(),
  xlab = NULL,
  ylab = NULL,
  
  axisLabels = c("show", "internal", "none"),
  columnLabels = colnames(numeric_steam),
  labeller = "label_value",
  switch = NULL,
  showStrips = NULL,
  legend = NULL,
  cardinality_threshold = 15,
  progress = NULL,
  proportions = NULL,
  #legends = deprecated()
)
```

\

There is only one strong correlation between variables num_reviews_total and peak_ccu. Num_reviews_total is the total number of reviews that the game has received. This consists of both positive and negative reviews. Peak_ccu is the games current highest peak concurrent player record. This is the greatest number of players playing the specific game at the same time. 

These two games have a strong, positive correlation of 0.84; as one variable increases, the other variable increases. This is visualized by the abline in the correlation matrix, which has a strong diagonal path to the right. 


### Taking a look at Price variable 

```{r,fig.show="hold", out.width="50%", echo=FALSE}

filtered_price_large <- small_steam %>% filter(price > 199.99)
#filtered_price_large[,c("name","price")]

filtered_price_small <- small_steam %>% filter(price < 199.99)
#filtered_price_small[,c("name","price")]

hist(filtered_price_small$price, breaks = 30,
     main = "Distribution of games that cost under $199.99",
     xlab = "Price of game in USD")
hist(filtered_price_large$price, breaks = 30,
     main = "Distribution of games that cost over $199.99",
     xlab = "Price of game in USD")

#ggplot(filtered_price_small, aes(x=price)) + 
#  geom_histogram(binwidth = 10) +
#  scale_x_continuous(breaks = seq(0, 199.99, by = 50))

#ggplot(filtered_price_large, aes(x=price)) +
#  geom_histogram(binwidth = 10) +
#  scale_x_continuous(breaks = seq(200, 1000, by = 50))

```
\

The variable price, denotes the base amount in USD that the game is priced at on Steam without any sales/discounts. The price variable shows a significant skew to the right where majority of the data points fall between 0-50. There are very few games that are priced above 199.99 USD, with there being three outliers that are all priced at 999.00 and 1000.00 USD. 
\

```{r, echo = FALSE}
filtered_price_large <- arrange(filtered_price_large, price)
pander(filtered_price_large[,c("name","price")])
```