---
title: "EDA_graphs"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the page for EDA and graphs. 

* Several summary tables and/or plots, each describing one, two or three variables in the data that you thought were interesting

* Explain the definitions of the variables in each table/plot

* Comment on the main features of each plot

* Include suitable labels and keys for each plot

* Make sure all plots would be readable if printed in black & white, and adjust the point sizes and/or line thicknesses to improve readability

* Lay out all tables so that they are clearly readable and clearly labelled, and do not use excessive significant figures

* All figures and tables should be numbered and should have captions


```{r, echo=FALSE, message = FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(data.table)
library(pander)
library(corrplot)

march_c <- read.csv("games_march2025_cleaned.csv")
small_steam <- march_c[c("name","price","dlc_count","about_the_game","short_description","windows","mac","linux",
                         "achievements","supported_languages","publishers","categories","genres",
                         "estimated_owners","average_playtime_forever","peak_ccu","num_reviews_total","pct_pos_total")]
```
\

#### Correlation matrix:

```{r, echo=FALSE}
library(corrplot)

numeric_steam <- small_steam[, sapply(small_steam, is.numeric)]
#summary(numeric_steam)

m <- cor(numeric_steam)
#corrplot(m, method = "number")
```
```{r, fig.show="hold", out.width="70%", echo=FALSE}
corrplot(m, method = "number")

pairs(numeric_steam[, 1:7], panel = function(x,y){
  points(x,y)
  abline(lm(y~x), col="red")})
```

