---
title: "Steam_test"
author: "Hannah Young"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the .rmd file for the project. Exploring the Steam data set. 

```{r, error = FALSE, warning=FALSE}
#install.packages("tidyverse")
#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages("data.table")
#install.packages("pander")

library(tidyverse)
library(ggplot2)
library(dplyr)
library(data.table)
library(pander)

march_c <- read.csv("games_march2025_cleaned.csv")
#march_c <- read.csv("C:/Users/Hannah Work/OneDrive/H_Laptop/Documents/games_march2025_cleaned.csv")

summary(march_c)
```
\

### Variables I want to include:

* appid
* name
* release_date
* required_age
* price
* dlc count
* about_the_game
\

*Note: I decided to keep about the game instead of the detailed and short descriptions.*
\

* header_image
* windows
* mac
* linux
* metacritic_score
* achievements
* recommendations
* supported_languages
* developers
* publishers
* categories
* genres
* user_score
* positive
* negative
* estimated_owners
* average_playtime_forever
* average_playtime_2weeks
* median_playtime_forever
* median_playtime_2weeks
* discount
* peak_ccu
* tags
\

*What do these mean: pct_pos_total, pct_pos_recent*
\

```{r}
nrow(march_c)
ncol(march_c)


```
\

![](C:/Users/Hannah Work/Downloads/cat.avif)
