---
title: "DATA301 - Group 4 Steam Project"
author: "Group 4"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the .rmd file for the project. Exploring the Steam data set. 

## Title

#### **Project title, names and IDs of students in your project group, date of submission**

• *State which dataset(s) your group worked on, and their source* **HANNAH**
\
<br>
We worked on the Steam Games Dataset 2025, and chose the games_march2025_cleaned subset to work on. This was sourced from Kaggle, where the data was created with Steam store page scraping and Steam Games Scraper repository, which gathers data using the Steam API and Steam Spy.

<br>
• *Explain briefly why the dataset is of interest, or what questions it could be used to answer; assume that the reader has never heard of your dataset* **HANNAH & COMBINED GROUP QUESTION**
\
<br>
This dataset provides extensive information about individual games that are available on the digital gaming distribution platform Steam. It was of interest for us because we thought that the game data could help to build informative analysis that tells a story and shows insightful correlations. We are able to conduct both broad and specific analysis which gives us alot of material to work with for answering specific questions. 

<br>
• *State the types of data in the dataset(s) and the structure of the dataset(s). Are the data numerical, categorical, or both? Time series? Coordinates? Diagnostic categories? This does NOT need to be an exhaustive list of every variable, just a few comments on the overall types.* **SAMUEL**

• *State how complete the dataset(s) are (i.e. how many missing, any structure to the missing data, whether there are errors in the data)* **SAMUEL**

• *If you used more than one dataset, state what steps you had to take to integrate the datasets* **HANNAH**
\
<br>
We only used the games_march2025_cleaned subset of the Steam Games Dataset 2025.

---

### **Background and Data**

#### Reduced variables:

* Name
* Release date
* Price
* DLC count
* About the game
* Short description
* Windows - Mac - Linux (combine)
* Achievements
* Supported languages
* Publishers
* Categories (selectively)
* Genres
* Estimated owners
* Average playtime forever
* peakCCU
* num_reviews total, pct pos total


--- 

### **Ethics, Privacy and Security**

• Brief discussion of any ethical considerations that apply to your project 

• Brief discussion of any privacy concerns that might arise connected to your project

• Brief discussion of what steps you could take to keep your project data and results secure (you do NOT need to carry this out, you just need to talk about it in the report). Make it clear what steps you actually did take vs. any you would take if given additional time and resources. 

---

### **Exploratory Data Analysis**

#### *Select a subset, highlighting features that you thought were interesting in the data. The plots do not have to be complicated; simple bar charts and scatter plots are fine.*


• Several summary tables and/or plots, each describing one, two or three variables in the data that you thought were interesting

• Explain the definitions of the variables in each table/plot

• Comment on the main features of each plot

• Include suitable labels and keys for each plot

• Make sure all plots would be readable if printed in black & white, and adjust the point sizes and/or line thicknesses to improve readability

• Lay out all tables so that they are clearly readable and clearly labelled, and do not use excessive significant figures

• All figures and tables should be numbered and should have captions

<br>

```{r, error = FALSE, warning=FALSE}
#install.packages("tidyverse")
#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages("data.table")
#install.packages("pander")

library(tidyverse)
library(ggplot2)
library(dplyr)
library(data.table)
library(pander)

march_c <- read.csv("games_march2025_cleaned.csv")
#march_c <- read.csv("C:/Users/Hannah Work/OneDrive/H_Laptop/Documents/games_march2025_cleaned.csv")

summary(march_c)

nrow(march_c)
ncol(march_c)
```

--- 

### **Individual Contributions**

State what contribution each member of the group made to the data preparation, the analysis and the report.

--- 

